---
title: "Mobile subscribers in East Africa"
output:
  html_document:
    theme: cerulean
---
## Disclaimer
This study is done as my personal study project. Any possible errors are highly regretted, I am still learning statistics.

##Source: World Development Indicators - World Bank

The dataset ca be found [here](http://data.un.org/Data.aspx?d=WDI&f=Indicator_Code%3aIT.CEL.SETS.P2)

##Load data and initial processing

```{r}
raw <- read.csv("mobile.csv", stringsAsFactors = FALSE)
mobile <- raw
mobile$Year <- as.numeric(mobile$Year)
mobile <- subset(mobile, !is.na(mobile$Year))
#drop Value.Footnotes
mobile <- mobile[,1:3]
mobile <- subset(mobile, Year>=2000)
mobile$Country.or.Area <- as.factor(mobile$Country.or.Area)
```

##Exploratory data analysis

We will focus on several regions:
###East Africa

```{r}
eAfrcountries <- c("Burundi", "Djibouti", "Eritrea", "Ethiopia", "Kenya", "Tanzania", "Uganda", "Rwanda")
eAfr <- subset(mobile, Country.or.Area %in% as.factor(eAfrcountries))

#drop unused country factor levels from other regions
eAfr$Country.or.Area = factor(eAfr$Country.or.Area)
```

##Average mobile subscriptions per country since 2000
```{r}
av_subscr <- tapply(eAfr$Value, eAfr$Country.or.Area, mean); av_subscr
```

##Kenya (per 100 people)
```{r, echo=FALSE}
Kenya <- subset(eAfr, Country.or.Area=="Kenya")
library(ggplot2)
ggplot(Kenya, aes(x=Year, y=Value)) + geom_bar(stat="identity", fill="lightblue", colour="black")
```

##Kenya compared to Tanzania (per 100 people)
```{r}
KenTan <- subset(eAfr, Country.or.Area %in% c("Kenya", "Tanzania"))
ggplot(KenTan, aes(x=Year, y=Value)) +
geom_bar(stat="identity", colour="black", fill="lightblue", position = "dodge") +
facet_grid(~ Country.or.Area)
  
KenTan <- subset(eAfr, Country.or.Area %in% c("Kenya", "Tanzania"))
ggplot(KenTan, aes(x=Year, y=Value, fill=Country.or.Area)) +
geom_bar(stat="identity", colour="blue", position = "dodge")
```

## East African countries (per 100 people)
```{r}
ggplot(eAfr, aes(x=Year, y=Value)) + geom_bar(stat="identity", fill="lightblue", colour="black") +
facet_wrap(~ Country.or.Area, ncol=2)
```


